### 2023-04-25 Qmd/paper.qmd
start paper

---
title: "Regression Discontinuity Design to Measure Incumbency Advantage in US Congressional Elections"
author: "Tyler Bagwell and Prayag Gordy"
format: pdf
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).